% =================================================================
\section{KPI Terms in the Schedule Objective}
\label{sec:kpi}
% =================================================================

Each KPI is written as a function of the schedule variables $x$
(with base camps fixed at $\hat b$ in Step~A) and then normalised by its
current-schedule value $\overline{K}_\bullet$ so that the weighted sum is
dimensionless. KPIs are oriented so that \emph{smaller is better};
maximise-type KPIs (5.1, 5.2, 5.3) enter as ``shortfall'' terms. Each
$\dist(\hat b_i,s)$, $\elev(s)$, $W_{mt}$, $q_t$ multiplying a binary
$x_{mts}$ is a known constant, so every term is linear after the
standard auxiliary-variable linearisations.

\paragraph{KPI 1.2 -- Intra-group travel dispersion.}
Per-team base-camp round-trip travel
$\mathrm{TD}_i=2\sum_{m\in M_i}\sum_{t,s}\dist(\hat b_i,s)\,x_{mts}.$
Group \emph{range} surrogate for the coefficient of variation: with
$\mathrm{TD}_g^{\max}\ge \mathrm{TD}_i\ge \mathrm{TD}_g^{\min}\ \forall
i\in g$,
\begin{equation}
  K_{1.2}=\sum_{g\in G}\bigl(\mathrm{TD}_g^{\max}-\mathrm{TD}_g^{\min}\bigr).
\end{equation}

\paragraph{KPI 1.3 -- Circadian shift cost.}
With base-camp time zone fixed, the perceived-night penalty
$\phi_{its}=\phi\!\bigl((\theta_t-[\tzone(s)-\tzone(\hat b_i)])\bmod 24\bigr)$
is a constant per (team, slot, stadium). Hence
\begin{equation}
  K_{1.3}=\sum_{i\in I}\sum_{m\in M_i}\sum_{t,s}\phi_{its}\,x_{mts}.
\end{equation}

\paragraph{KPI 1.4 -- Match-venue geographic dispersion.}
Let $a_{ic}\in\{0,1\}$ indicate team $i$ visits cluster $c$:
$a_{ic}\ge x_{mts}\ \forall m\in M_i,\,s:\cl(s)=c$. Then
\begin{equation}
  K_{1.4}=\sum_{i\in I}\sum_{c\in\{\mathrm W,\mathrm C,\mathrm E\}} a_{ic}.
\end{equation}

\paragraph{KPI 1.6 -- Rest asymmetry between opponents.}
Let $r_{mi}$ be team $i$'s rest hours before match $m$ (linear in $x$
via slot times). With $\Delta_m\ge r_{m,i(m)}-r_{m,j(m)}$ and
$\Delta_m\ge r_{m,j(m)}-r_{m,i(m)}$,
\begin{equation}
  K_{1.6}=\sum_{m\in M}\Delta_m .
\end{equation}

\paragraph{KPI 1.7 -- Entry and visa restriction exposure.}
Severity-weighted count of affected teams playing at US stadiums:
\begin{equation}
  K_{1.7}=\sum_{i\in I}\sigma_i
          \sum_{m\in M_i}\sum_{t}\sum_{s\in S^{\mathrm{US}}} x_{mts}.
\end{equation}

\paragraph{KPI 2.2 -- Per-team heat load.}
With excess WBGT $h_{mt}=\max(0,W_{mt}-28)$ a constant,
\begin{equation}
  K_{2.2}=\sum_{i\in I}\sum_{m\in M_i}\sum_{t,s} h_{mt}\,x_{mts}.
\end{equation}

\paragraph{KPI 3.3 -- Round-order balance (first-mover).}
Let $a_i=\sum_{r\in\{1,2\}}\mathrm{pos}(i,r)$ count earlier-kickoff
appearances in the two sequential rounds ($\mathrm{pos}(i,r)$ linear in
$x$ and the round-slot ordering). Balancing targets $a_i=1$; with
$e_i\ge a_i-1$, $e_i\ge 1-a_i$,
\begin{equation}
  K_{3.3}=\sum_{i\in I} e_i .
\end{equation}

\paragraph{KPI 4.1 -- Venue-load balance.}
Let $n_s=\sum_{m,t} x_{mts}$ and mean $\bar n=|M|/|S|$. Mean-absolute-
deviation surrogate for the CV with $d_s\ge n_s-\bar n$,
$d_s\ge \bar n-n_s$:
\begin{equation}
  K_{4.1}=\sum_{s\in S} d_s .
\end{equation}

\paragraph{KPI 4.2 -- Same-city overlap.}
With $o_{ct}\ge \sum_{m}\sum_{s:\,\mathrm{city}(s)=c} x_{mts}-1$,
$o_{ct}\ge 0$ counting simultaneous same-city matches,
\begin{equation}
  K_{4.2}=\sum_{c}\sum_{t} o_{ct}.
\end{equation}


\paragraph{KPI 5.2 -- Marquee-match slot quality (maximise) with overlap.}
\begin{equation}
  K_{5.2}=\overline{\mathrm{MSQ}}
          -\sum_{m\in M}\sum_{t,s}\mu_m\,q_t\,x_{mts}
          +\lambda\!\!\sum_{\{m,m'\}\subseteq P}\!\! o^P_{mm'},
\end{equation}
where $o^P_{mm'}\ge \sum_s x_{mts}+\sum_s x_{m'ts}-1$ at each common slot
$t$ penalises two high-profile matches sharing a slot.

\paragraph{KPI 5.3 -- Host-city economic equity (maximise).}
Venue commercial value $\mathrm{VC}_s=\sum_{m,t}\mu_m q_t x_{mts}$,
mean $\overline{\mathrm{VC}}=\tfrac1{|S|}\sum_s\mathrm{VC}_s$.
Gini-replacing mean-absolute-deviation dispersion with
$p_s\ge \mathrm{VC}_s-\overline{\mathrm{VC}}$,
$p_s\ge \overline{\mathrm{VC}}-\mathrm{VC}_s$:
\begin{equation}
  K_{5.3}=\sum_{s\in S} p_s .
\end{equation}

% 